---
title: "Reproducible research assignment"
author: "Khandaker Tanveer Ahmed"
date: "5/12/2020"
output: html_document
---

---
title: 'Reproducible research: assignment 1'
author: "Khandaker Tanveer Ahmed"
date: "5/12/2020"
output: html_document
---
## Getting the data
```{r}
stepdata <- read.csv("activity.csv", header = TRUE)
head(stepdata)
```

## Total steps taken a day
```{r steps}
library(magrittr)
library(dplyr)
databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps)) %>%na.omit()
hist(databydate$tsteps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
```

## Mean and median of total steps a day
```{r}
mean(databydate$tsteps)

median(databydate$tsteps)
```

## Time series plot for total steps
```{r time series}
library(ggplot2)
databyinterval <- stepdata%>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps)) 
ggplot(databyinterval, aes(x=interval, y=tsteps))+ geom_line()
```

## Interval checking
```{r interval}
databyinterval[which(databyinterval$tsteps== max(databyinterval$tsteps)),]
```

# Imputing missing values
## Total missing values
```{r}
missingVals <- sum(is.na(data))
missingVals
```

## Replacing missing values
### I will use the mean for that 5 -minute interval to replace all the missing values in the dataset. At the end, I will check if all the NAs have been replaced
```{r}
library(magrittr)
library(dplyr)

replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```

## The mean and median total number of steps taken per day.
```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
```

## Summary of new data : mean & median
```{r}
summary(FullSummedDataByDay)
```

## Making a histogram
```{r}
hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```

## Compare the mean and median of Old and New data
```{r}
oldmean <- mean(databydate$tsteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)
```

## Old mean and New mean
```{r}
oldmean

newmean
oldmedian <- median(databydate$tsteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)

oldmedian

newmedian
```

## Differences in activity patterns between weekdays and weekends
```{r differences}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )
library(ggplot2)
meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")

ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```